---
title: "Homework_6"
author: "Vaishali Verma"
date: "2023-06-20"
output: html_document
---

#### Libraries
##### My last name is Verma so I will analyze heights
```{r libraries, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
```

#### Getting data, splitting data, subsetting data

```{r data}
data <- gcookbook::heightweight

subset_male <- subset(data, data$sex == 'm')
subset_female <- subset(data, data$sex == 'f')

set.seed(100)
random_sample_female <- subset_female[sample(nrow(subset_female), 15, replace = FALSE), ]
random_sample_male <- subset_male[sample(nrow(subset_male), 15, replace = FALSE), ]
random_sample_female_summary <- summary(random_sample_female)
random_sample_male_summary <- summary(random_sample_male)

```
#### Confidence in two means being same or different
##### Using t-test to test means
###### H0 - there is no significant difference between the mean of two groups
###### H1 - there is a significant difference between the mean of two groups
```{r t_test}
t_test_result <- t.test(random_sample_female$heightIn, random_sample_male$heightIn, alpha = 0.05)
t_test_result
p_value <- t_test_result$p.value
p_value
my_confidence <- ifelse(p_value < 0.05, "reject", "fail to reject")
my_confidence
```
###### Thus, I fail to reject H0
###### There is not enough evidence to say the means are different

#### Single Random Dataframe
```{r combine_random_dataframe}
combined_data <- bind_rows(random_sample_female, random_sample_male)
mean_female_random <- mean(random_sample_female$heightIn)
mean_male_random <- mean(random_sample_male$heightIn)

#confidence intervals
female_t_test <- t.test(random_sample_female$heightIn, alpha= 0.05)
male_t_test <- t.test(random_sample_male$heightIn, alpha=0.05)

# Create a jittered points chart
plot <-  ggplot(combined_data, aes(x = sex, y = heightIn)) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.5) 
plot <- plot + geom_errorbar(data = random_sample_male, aes(x = sex, y = heightIn, ymin = male_t_test$conf.int[1], ymax = male_t_test$conf.int[2]),
                width = 0.2, color = "blue") 
plot <-  plot + geom_errorbar(data = random_sample_female, aes(x = sex, y = heightIn, ymin = female_t_test$conf.int[1], ymax = female_t_test$conf.int[2]),
                       width = 0.2, color = 'red')
plot <- plot + geom_hline(data = random_sample_female, aes(yintercept = mean_female_random), color = 'pink') +
  geom_label(aes(x = 'f', y= 75), color ='pink', label = paste("Sample Mean-", round(mean_female_random, 2)))
plot <- plot + geom_hline(data = random_sample_male, aes(yintercept = mean_male_random), color = 'lightblue') +
  geom_label(aes(x = 'm', y=75), color = 'lightblue', label = paste("Sample Mean-", round(mean_male_random, 2)))
plot
```

#### Perform t-test for grades 6 to 10, p-value and reject or fail-to-reject
```{r grade_t_test}
data$grade<-as.factor(round(data$ageYear) -6)

#t tests
t_test_grade6 <- t.test(subset(data, data$grade == 6 & data$sex == 'm')$heightIn, subset(data, data$grade == 6 & data$sex == 'f')$heightIn, alpha = 0.05)
t_test_grade7 <- t.test(subset(data, data$grade == 7 & data$sex == 'm')$heightIn, subset(data, data$grade == 7 & data$sex == 'f')$heightIn)
t_test_grade8 <- t.test(subset(data, data$grade == 8 & data$sex == 'm')$heightIn, subset(data, data$grade == 8 & data$sex == 'f')$heightIn)
t_test_grade9 <- t.test(subset(data, data$grade == 9 & data$sex == 'm')$heightIn, subset(data, data$grade == 9 & data$sex == 'f')$heightIn)
t_test_grade10 <- t.test(subset(data, data$grade == 10 & data$sex == 'm')$heightIn, subset(data, data$grade == 10 & data$sex == 'f')$heightIn)

grade <- c("Grade_6", "Grade_7", "Grade_8", "Grade_9", "Grade_10")
alpha <- rep(0.05, 5)
p_value <- c(t_test_grade6$p.value, t_test_grade7$p.value, t_test_grade8$p.value, t_test_grade9$p.value, t_test_grade10$p.value)
p_value
result <- c()
for (i in 1:length(p_value)) {
  if (p_value[i] > alpha[i]) {
    result <- c(result, "fail to reject")
  } else {
    result <- c(result, "reject")
  }
}
test_table <- data.frame(Name = grade, Alpha = alpha, pValue = p_value, Result=result)
test_table
```

#### Optional - facet grid to show mean height and 95% CI for male and female
```{r facet_grid_mean_ht_ci, message=FALSE, warning=FALSE}
ggplot(data, aes(x = sex, y = heightIn)) +
  geom_point(width = 0.2, height = 0, alpha = 0.3) +
  stat_summary(
    fun = mean, 
    geom = "errorbar", 
    position = position_dodge(width = 0.2),
    width = 0.2,
    fun.data = mean_cl_normal,
    na.rm = TRUE,
    color = "red"
  ) +
  facet_wrap(~ grade, nrow = 2)
```